---
title: "Effectiveness of chlorine on the microbial reduction in agricultural water"
author: Naim Montazeri, Simon Riley, and Arie Havelaar
format:
  docx:
    toc: true
    number-sections: true
    highlight-style: github
code-wrap: true
editor: visual
Date: Jan 20, 2025
---

# Preliminary setup

Run this script from within the R project - the code now uses relative paths which may not work otherwise.

**Load Required Packages**

```{r}
#| warning: false
#| message: false

# Load packages
library(here)
library(writexl)
library(tidyverse)
library(ggpubr)
library(nlme)
library(ggResidpanel)
library(performance)
library(emmeans)
library(marginaleffects)
library(multcomp)
library(multcompView)

sessioninfo::session_info(pkgs = c("attached"))
```

**Set global options**

```{r}
options(contrasts = c('contr.sum', 'contr.poly'), 
        dplyr.width = Inf, pillar.print_max = 50, pillar.print_min = 50)

knitr::opts_chunk$set(eval = T, echo = T, cache = T, fig.align = 'center')

set.seed(1484) #accroding to https://www.random.org
```

**Define custom functions**

The candidate models are Weibull and log-logistic. The derivation of the reparameterized log-logistic is provided at the end of this document.

```{r}
# Weibull and reparameterized weibull
 wb = function (x, log10n0, alpha, beta) {        
   log10n0 - (x/exp(alpha))^beta
 }
 
 wbrepar = function(x, log10n0, TD, beta, D = 1){ 
  log10n0 - D*(x/TD)^beta
}

# Log-logistic and reparameterized log-logistic 
lg = function(x, log10n0, kappa, sigma){          
   log10n0 - log10(1+ exp((log(x)-kappa)/sigma^2))
 }
 
lgrepar = function(x, log10n0, TD, sigma, D = 1){  
  log10n0 - log10(1+ exp((log(x)- (log(TD) - log(10^D - 1)*sigma^2))/sigma^2))
}
```

# Chlorine demand (CLD)

## Data entry

```{r}
chl = read.delim(here('Data (study 1)', "obj1_chl_NM.txt")) |> 
  mutate(across(1:2, factor), 
         sample = fct_relevel(sample, 'di', 'ag1', 'ag2', 'ag3', 'ag4'))
str(chl)

saveRDS(chl, here('Results (study 1)', 'CLD_all_ave.RDS'))

# Grouping based on sanitizer for each Ag water
CLD_m1 = readRDS(here('Results (study 1)', "CLD_all_ave.RDS")) |>
  dplyr::filter(sample=="ag1" | sample=="ag2" |
                  sample=="di" &  sanitizer=="accutab")
```

## Plot CLD

```{r}
CLD_m1 =  readRDS(here('Results (study 1)', "CLD_all_ave.RDS")) |>
  dplyr::filter(sample=="ag1" | sample=="ag2" |
                  sample=="di" &  sanitizer=="accutab") |> 
  mutate(chl.added = factor(chl.added))

str(CLD_m1)

var_names = c(
 "sample" = "Samples",
 "ag1"="Ag1",
 "ag2" = "Ag2",
 "di" = "DI water"
  )

p_chl_demand = CLD_m1 |>
  filter (sample =="di"|sample=="ag1" | sample=="ag2") |>
  ggplot (aes(x=chl.added, y=CLD.ave, fill=sample)) +
  geom_bar (position = position_dodge(0.8),
            width=0.8,
            color ="black", 
            stat="identity",
            linewidth=0.3) +
  facet_grid (~ sample, labeller = as_labeller(var_names)) +
  geom_errorbar(aes(ymin=CLD.ave-CLD.se, 
                    max=CLD.ave+CLD.se),
                position = position_dodge(0.8),
                linewidth=0.3,
                width=0.2) +
  labs (
    x = "Free available chlorine dose (ppm)", 
    y = "Chlorine demand (ppm)", 
    size = 9, face="bold") +
  scale_fill_manual(values=c("grey60",  "grey60", "grey60"),
                    name="",
                    breaks=c("di", "ag1", "ag2"),
                    labels=c("DI water", "Ag1", "Ag2")) +
  ylim (-0.1, 8) +
  theme_bw (base_family="") +
  theme(
    text = element_text(family = "Arial"),
    legend.position = "none", 
    plot.title = element_text(color="black", size=9, face = "bold", hjust = 0.5),
    axis.text.x=element_text(size = 7, color="black", face="bold", hjust = 0.5, vjust = 0.5),
    axis.text.y=element_text (size = 7, color="black", face="bold"),
    axis.title.x = element_text (size = 9, vjust = -1),
    axis.title.y = element_text (size = 9, vjust = 2),
    panel.grid.major = element_line (color="gray", linewidth=0.1, linetype="longdash"),
    panel.grid.minor = element_line (color="lightgray", linewidth=0.1, linetype="dotted"),
    panel.background = element_rect (fill =  "transparent"),
    legend.background = element_rect (fill = "transparent", color = "transparent"),
    plot.background = element_rect (fill = "transparent", color = "transparent")) +
  guides(
    color = guide_legend(override.aes = list(size = 5))
  )

p_chl_demand

ggsave(file=here('Results (study 1)', "Fig_1 CLD_plot.tiff"),
       plot = p_chl_demand, 
       width=9, height=3.9, units = c("cm"), dpi=600)
```

# Control group

## Data entry and basic calculations

```{r}
m1c = readRDS(here('Data (study 1)', 'obj1_m1.rda')) |> 
 filter(sample == "di" & time == 5) |>
 mutate(across(c(1, 4, 7:8, 10), factor))

m1c_w = m1c |>
 tidyr::spread (rep, survival)

# calculate mean, sd, and se
m1c_w$mean = apply(
 m1c_w[,c(11:12)], 1, base::mean, na.rm=TRUE) # calculate mean
m1c_w$sd = apply( 
  m1c_w[,c(11:12)], 1, stats::sd, na.rm=TRUE) # calculate Std Dev
m1c_w$se = m1c_w$sd/sqrt(3) # calculate Std Error

m1c_w |>
 dplyr::select (sanitizer, sample, chl.measured, temp, time, microb, assay, mean, sd, se) |>
 saveRDS(here('Results (study 1)', "m1c_w.RDS"))

# Log10 reduction for E coli 0 ppm to 2 ppm
ec_chl0 <- m1c |>
  filter(microb == "ec" & chl.measured == 0 & time=="5") |>
  dplyr::select(survival)
mean(ec_chl0$survival)
sd(ec_chl0$survival)/sqrt(3)

ec_chl2 <- m1c |>
  filter(microb == "ec" & chl.measured == 1.8 & time=="5") |>
  dplyr::select(survival)
mean(ec_chl2$survival)
sd(ec_chl2$survival)/sqrt(3)

mean(ec_chl0$survival-ec_chl2$survival)
sd(ec_chl0$survival-ec_chl2$survival)/sqrt(3)

# Log reduction for TV 0 ppm to 4 ppm
tv_chl0 <- m1c |>
  filter(microb == "tv" & chl.measured == 0 & time=="5") |>
  dplyr::select(survival)
mean(tv_chl0$survival)
sd(tv_chl0$survival)/sqrt(3)

tv_chl2 <- m1c |>
  filter(microb == "tv" & chl.measured == 1.8 & time=="5") |>
  dplyr::select(survival)
mean(tv_chl2$survival)
sd(tv_chl2$survival)/sqrt(3)

mean(tv_chl0$survival-tv_chl2$survival)
sd(tv_chl0$survival-tv_chl2$survival)/sqrt(3)

tv_chl4.1 <- m1c |>
  filter(microb == "tv" & chl.measured == 4.1 & time=="5") |>
  dplyr::select(survival)
mean(tv_chl4.1$survival)
sd(tv_chl4.1$survival)/sqrt(3)

mean(tv_chl0$survival-tv_chl4.1$survival)
sd(tv_chl0$survival-tv_chl4.1$survival)/sqrt(3)

# Max inactivation
mean(ec_chl0$survival) - 1
mean(tv_chl0$survival) - 0.347
```

## Plot control group

```{r}
m1c_w = readRDS(here('Results (study 1)', "m1c_w.RDS")) |>
 filter(time=="5")
m1c_w$chl.measured.fac = as.factor (round(m1c_w$chl.measured, 0))

# LOD for each microbe
lod_tv = 0.347
lod_ec = 1.0

p_tv = m1c_w |>
 filter(microb=="tv") |>
 ggplot (aes(x=chl.measured.fac, y=mean)) + 
 geom_bar(stat = "identity", position = position_dodge(0.1), 
         linewidth = 0.3, fill = "darkgray", color = 'black') +
 geom_errorbar (aes(x=chl.measured.fac, ymin=mean-se, ymax=mean+se), 
                width=0.3, alpha=0.5, color = "black",
                position = position_dodge(0.1)) +
 geom_hline (yintercept=lod_tv, linetype="dashed", color = "#00529b", linewidth = 0.3) +
 scale_shape_manual(values=c (3, 6)) +
 ggtitle (expression(paste("Tulane virus"))) +
 labs (y= bquote (Log[10]~PFU~per~ml)) +
 ylim (0, 7) +
  theme_bw (base_family="") +
  theme(
    plot.title = element_text(color="black", size=9, face = "bold", hjust = 0.5),
    axis.text.x=element_text (size = 9, color="black", face="bold"),
    axis.text.y=element_text (size = 9, color="black", face="bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text (size = 9, vjust = 2),
    panel.grid.major = element_line (color="gray", linewidth=0.1, linetype="longdash"),
    panel.grid.minor = element_line (color="lightgray", linewidth=0.1, linetype="dotted"),
    panel.background = element_rect (fill =  "transparent"),
    legend.background = element_rect (fill = "transparent", color = "transparent"),
    plot.background = element_rect (fill = "transparent", color = "transparent"))

 p_ec = m1c_w |>
 filter(microb=="ec" & chl.measured!="14.3") |>
 ggplot (aes(x=chl.measured.fac, y=mean)) + 
 geom_bar(stat = "identity", position = position_dodge(0.1), 
         linewidth = 0.2, fill = "darkgray", color = 'black') +
 geom_errorbar (aes(x=chl.measured.fac, ymin=mean-se, ymax=mean+se), 
                width=0.3, alpha=0.5, color = "black",
                position = position_dodge(0.1)) +
 geom_hline (yintercept=lod_ec, linetype="dashed", color = "#00529b", linewidth = 0.3) +
 scale_shape_manual(values=c (3, 6)) +
 ggtitle (expression(italic("E. coli"))) +
 labs (y= bquote (Log[10]~CFU~per~ml)) +
 ylim (0, 7) +
  theme_bw (base_family="") +
  theme(
    plot.title = element_text(color="black", size=9, face = "bold", hjust = 0.5),
    axis.text.x=element_text (size = 9, color="black", face="bold"),
    axis.text.y=element_text (size = 9, color="black", face="bold"),
    axis.title.x = element_blank(),
    axis.title.y = element_text (size = 9, vjust = 2),
    panel.grid.major = element_line (color="gray", linewidth=0.1, linetype="longdash"),
    panel.grid.minor = element_line (color="lightgray", linewidth=0.1, linetype="dotted"),
    panel.background = element_rect (fill =  "transparent"),
    legend.background = element_rect (fill = "transparent", color = "transparent"),
    plot.background = element_rect (fill = "transparent", color = "transparent"))

combined_plot = ggpubr::ggarrange(p_ec,
                                  p_tv, ncol = 2,
                                   labels = c("A", "B"),
                                   font.label = list(size=9, face="bold", family="Arial" 
                                   ))
combined_plot <- annotate_figure(combined_plot, 
                                 bottom = text_grob("Free available chlorine (ppm)",
                                                    size = 9,
                                                    family = "Arial"))
ggsave(plot=combined_plot,
       file= here('Results (study 1)', "Fig_2_control_group_plot.tiff"), 
       width=9, height=5, units = c("cm"), dpi=600)
```

# Modeling microbial inactivation kinetics

## Data entry and approach

```{r}
m1_file = here('Data (study 1)', 'obj1_m1_NM.txt')
m1 = read.table(file = m1_file, sep = '\t', header = T) |> 
  mutate(across(c(1, 4, 7:8, 10), factor), 
         temp_fac = factor(temp), 
         time_fac = factor(time, labels = c('5min', '10min')), 
         microb = fct_recode(microb, 'ec' = 'ecoli', 'tv' = 'tv'))

saveRDS(m1, here('Data (study 1)', 'obj1_m1.rda'))
```

Since there were insufficient time points included in the study to fit a primary model, time was initially included as a (categorical) fixed effect for the model parameters in the Weibull and log-logistic models. In all cases, however, neither time nor any interactions with time were statistically significant (p \> 0.05), and thus it was determined that - given the impossibility of fitting a primary model and the clear indication that the overwhelming majority of inactivation was occurring, for all treatments and microbes, within the first 5 minutes - to refit the model using only the data from the 5min exposure.

## Model Development and Evaluation

```{r}
m1 = readRDS(here('Data (study 1)', 'obj1_m1.rda')) |> 
  filter(sample != 'di' & time == 5) |> 
  droplevels()

m1_mod_wb = gnls(survival ~ wb(x = chl.measured, 
                               log10n0 = log10n0, 
                               alpha = alpha,
                               beta = beta), 
                 data = m1, 
                 params = log10n0+alpha+beta ~ sample*microb, 
                 start = list(log10n0 = c(6, rep(0, 3)), 
                              alpha = c(5, rep(0, 3)), 
                              beta = c(1, rep(0, 3))))

m1_mod_lg = gnls(survival ~ lg(x = chl.measured, 
                               log10n0 = log10n0,
                               kappa = kappa, 
                               sigma = sigma), 
                 data = m1, 
                 params = log10n0+kappa+sigma ~ sample*microb, 
                 start = list(log10n0 = c(5, rep(0, 3)), 
                               kappa = c(.1, rep(0, 3)), 
                               sigma  = c(.2, rep(0, 3))))
```

Initial comparison via AICc indicates log-logistic is the best fitting:

```{r}
#| warning: false
compare_performance(m1_mod_wb, m1_mod_lg, metrics = c('AICc'))
```

Examination of the residuals suggests some evidence of heteroscedasticity (although likely actually the result of censoring of observations below LOD). Nevertheless, the model is compared with heteroscedasticity for any improvement.

```{r}
#| warning: false

m1_res_lg = resid(m1_mod_lg, type = 'normalized')
m1_fit_lg = fitted(m1_mod_lg)
resid_auxpanel(m1_res_lg, m1_fit_lg)
```

Power model is best fitting.

```{r}
#| warning: false
m1_mod_lg_exp = update(m1_mod_lg, weights = varExp())
m1_mod_lg_pow = update(m1_mod_lg, weights = varPower(), 
                    control = gnlsControl(maxIter = 100))
m1_mod_lg_ident = update(m1_mod_lg, weights = varIdent(form = ~1|microb))

compare_performance(m1_mod_lg, m1_mod_lg_exp, m1_mod_lg_pow,
                    m1_mod_lg_ident, metrics = c('AICc')) |> 
  dplyr::select(Name, Model, AICc, AICc_wt)
```

Refit the model using the re-parameterized version of the model. The code below converts the estimates in the original model into the corresponding estimates in the reparameterized model, so that they can be used as starting values.

```{r}
# Extract design matrix for each treatment combination and name the rows
# accordingly
E = model.matrix(~ sample*microb, data = m1) |> 
  unique()
ids = as.numeric(rownames(E))
rownames(E) = t(apply(m1[ids, c('sample', 'microb')], 1, paste0, collapse = ' '))

# Get ordering for marginal means and make sure the rows in the "E" matrix match
#emmeans(m1_mod_lg, ~ sample:microb, param = 'kappa')
E = E[c(3, 4, 1, 2),] 

# calculate marginal means for each treatment combination
kappa_emms = emmeans(m1_mod_lg_pow, ~ sample:microb, param = 'kappa')|> 
    as.data.frame() |> 
    dplyr::select(emmean) |> 
    unlist()

n0_emms = emmeans(m1_mod_lg_pow, ~sample:microb, param = 'log10n0')|> 
    as.data.frame() |> 
    dplyr::select(emmean) |> 
    unlist()

sigma_emms = emmeans(m1_mod_lg_pow, ~sample:microb, param = 'sigma') |> 
    as.data.frame() |> 
    dplyr::select(emmean) |> 
    unlist()

# Calculate the time to 3 log10 reductions implied by the fitted estimates
# for each treatment combination
td_emms = log(10^3-1)*sigma_emms + kappa_emms

# Then, calculate the coefficients (based on the E matrix) implied by those
# 3 log10 reductions
td_coefs = round(solve(E)%*%td_emms, 1)
n0_coefs = round(solve(E)%*%n0_emms, 1)
sigma_coefs = round(solve(E)%*%sigma_emms, 1)

# These values should be extremely close to the fitted values in the
# reparameterized model
```

```{r}
m1_mod = gnls(survival ~ lgrepar(x = chl.measured, 
                                 log10n0 = log10n0, 
                                 TD = TD, 
                                 sigma = sigma, 
                                 D = 3), 
              data = m1, 
              params = log10n0+TD+sigma ~ sample*microb, 
              weights = varPower(), 
              start = list(log10n0 = n0_coefs, 
                            TD  = td_coefs, 
                            sigma  = sigma_coefs))


saveRDS(m1_mod, here('Results (study 1)', "m1_mod.RDS"))

# Extract estimates and their 95% CI
m1_mod_coef = coef(m1_mod)
m1_mod_confint = confint(m1_mod)

as.data.frame (cbind(
  Coefs = names(m1_mod_coef),
  Est = round (as.numeric(m1_mod_coef), 3),
  lwr = round (m1_mod_confint[,1], 3),
  upr = round (m1_mod_confint[,2], 3)
  )) |>
 write_xlsx(here('Results (study 1)', "m1_mod_summary.xlsx"))

readxl::read_xlsx(here('Results (study 1)', "m1_mod_summary.xlsx"))
```

## Estimation, Inference, and Plotting

Only microbe species has a significant effect on the initial population (log10n0) parameter, while there is a significant interaction effect between microbe and sample on the chlorine required for log 3 reduction (TD parameter), and for the sigma parameter (which characterizes the variation in chlorine susceptibility within the microbial population).

```{r}
m1_mod = readRDS(here('Results (study 1)', "m1_mod.RDS"))

(m1_ftest_n0 = joint_tests(m1_mod, param = 'log10n0'))
(m1_ftest_TD = joint_tests(m1_mod, param = 'TD'))
(m1_ftest_sigma = joint_tests(m1_mod, param = 'sigma'))

write_xlsx(list(log10n0 = as.data.frame(m1_ftest_n0), 
                TD = as.data.frame(m1_ftest_TD), 
                sigma = as.data.frame(m1_ftest_sigma)), 
           path = here('Results (study 1)', 
                       'm1_param_ftests.xlsx'))
```

Estimates and contrasts for the different parameters:\*\*\*

```{r}
#| warning: false
(m1_emm_n0 = emmeans(m1_mod, pairwise ~ microb, param = 'log10n0'))
(m1_emm_TD = emmeans(m1_mod, pairwise ~ microb:sample, param = 'TD'))
(m1_emm_sigma = emmeans(m1_mod, pairwise ~ microb:sample, param = 'sigma'))
```

## Plot 3-log10 reduction (TD)

```{r}
#| fig-width: 4
#| fig-height: 4
#| warning: false

m1_emm2_TD = emmeans(m1_mod, pairwise ~ sample:microb,
                        param = 'TD')

sample_labs = c(ag1 = 'Ag1', ag2 = 'Ag2')

m1_ch3d_plot = m1_emm2_TD |> 
  cld(Letters = letters, reversed = T, adjust = 'none') |> 
  as.data.frame() |> 
  mutate(.group = gsub(' ', '', .group), 
         Microbe = factor(microb, labels = c('E. coli', 'Tulane Virus')) ) |> 
  ggplot(aes(x = Microbe, fill = Microbe))+
  facet_grid(cols = vars(sample), labeller = as_labeller(sample_labs))+
  geom_col(aes(y = emmean), show.legend = F, color = "black", linewidth = 0.2, width = 0.4) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2, alpha=0.4)+
  geom_text(aes(y = upper.CL+2, label = .group)) +
  scale_y_continuous('Free Available Chlorine (ppm)')+
  scale_x_discrete('Microbes', labels = c(expression(italic("E. coli")), "Tulane Virus"))+
  scale_fill_manual(values = c("E. coli" = "gray80", "Tulane Virus" = "gray60")) +
 theme(
  plot.title = element_text(color="black", size=8, face = "bold", hjust = 0.5),
  axis.text.x=element_text(size = 8, color="black", face="bold", hjust = 0.5, vjust = 0.5),
  axis.text.y=element_text (size = 8, color="black", face="bold"),
  axis.title.x = element_text (size = 8, vjust = -1),
  axis.title.y = element_text (size = 8, vjust = 2),
  panel.grid.major = element_line (color="gray", linewidth=0.1)
 )

m1_ch3d_plot

ggsave(plot=m1_ch3d_plot, file=here('Results (study 1)', "m1_ch3d_plot.tiff"), 
       width=14, height=7, units = c("cm"), dpi=600)
```

## Plot microbial inactivation

```{r}
#| warning: false

m1_grid = expand.grid(sample = levels(m1$sample), 
                      microb = levels(m1$microb),
                      chl.measured = seq(0, 36, 0.5)) |> 
  filter(!(microb == 'ec' & chl.measured > 13.5))

m1_preds = avg_predictions(m1_mod, newdata = m1_grid, 
                           by = c('sample', 'microb', 'chl.measured'))

microb_labs = c(ec = expression(italic('E. coli')), tv = 'Tulane Virus')

p_m1= m1_preds |>
  mutate(microb = factor(microb, labels = c('italic(E.~coli)', 'Tulane~Virus')), 
         LOD = ifelse(microb == 'Tulane~Virus', 0.36, 1)) |> 
  ggplot(aes(x = chl.measured)) +
  facet_grid(cols = vars(microb), labeller = label_parsed, scales = 'free_x') +
  geom_hline(aes(yintercept = LOD), color = 'black', linetype = 3)+
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, group = sample), alpha = .2, 
              fill = 'grey45') +
  geom_line(aes(y = estimate, linetype = sample), color = "black") +
  geom_point(aes(y = survival, shape = sample),
             data = mutate(m1, microb = factor(microb, 
                                               labels = c('italic(E.~coli)', 
                                                          'Tulane~Virus'))),
             size = 1.5) +
  guides (fill = guide_legend(title = 'Sample:'),
         color = guide_legend(title = 'Sample:'), 
         linetype = guide_legend(title = 'Sample:'),
         shape = guide_legend(title = 'Sample:')) +
  labs(x = "Free available chlorine (ppm)", 
       y = bquote(Log[10]~microbial~count~per~ml)) +
  scale_fill_manual(labels = c('Ag1', 'Ag2')) +
  scale_shape_manual(values = c(1, 17), 
                     labels = c('Ag1', 'Ag2')) +
  scale_linetype_manual(values = c("dashed", "solid"), 
                        labels = c('Ag1', 'Ag2')) +
  theme(text = element_text(family = "Arial"),
        legend.position = 'inside', legend.position.inside = c(.88, .8),
        legend.direction = "vertical", 
        legend.text = element_text(size = 7),
        legend.title = element_text(size = 7, face = "bold"),
        legend.key.height = unit(0.4, "cm"),
        plot.title = element_text(color = "black", size = 9, face = "bold", hjust = 0.5),
        axis.text.x = element_text(size = 9, color = "black", face = "bold", hjust = 0.5, vjust = 0.5),
        axis.text.y = element_text(size = 9, color = "black", face = "bold"),
        axis.title.x = element_text(size = 9, vjust = -1),
        axis.title.y = element_text(size = 9, vjust = 2),
        panel.grid.major = element_line(
          color = "gray", linewidth = 0.1, linetype = "longdash"),
        panel.grid.minor = element_line(
          color = "lightgray", linewidth = 0.1, linetype = "dotted"),
        panel.background = element_rect(fill = "transparent"),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        plot.background = element_rect(fill = "transparent", color = "transparent"))

p_m1

ggsave(file = here('Results (study 1)', "Fig_3_trend_plot.tiff"), 
       plot = p_m1,
       width=14, height=7, units = c("cm"), dpi=600)
```

These are model-based estimates of mean (along with their standard error) of the log10 PFU for each treatment combination.

```{r}
m1_tab_grid = m1 |> 
  dplyr::select(sample, microb, chl.added) |> 
  rename(chl.measured = chl.added) |> 
  distinct()
m1_ests = avg_predictions(m1_mod, newdata = m1_tab_grid, 
                          by = c('microb', 'sample', 'chl.measured'))
m1_ests_tab = m1_ests |> 
  as.data.frame() |> 
  rename(Sample = sample, Microbe = microb) |> 
  mutate(across(4:5, as.numeric), 
         estimate = paste0(sprintf("%.2f", estimate),
                           ' \u00b1 ', sprintf("%.2f", std.error)), 
         chl.measured = paste0(chl.measured, ' ppm')) |> 
  dplyr::select(Microbe, Sample, chl.measured, estimate) |> 
  pivot_wider(names_from = chl.measured, values_from = estimate, 
              names_prefix = '')
print(m1_ests_tab)

write_xlsx(m1_ests_tab, 
           path = here('Results (study 1)',
                       'Estimated Conc. for 3 Log Reduction.xlsx'))
```

# Reparameterizing the Log-Logistic Model

$$log_{10}(N_t) = log_{10}(N_0) - log_{10}(1+e^{\frac{log(x) - \kappa}{\sigma^2}})$$ Letting $C_3$ be the concentration at which $log_{10}(N_t)$ is 3 below $log_{10}(N_0)$, it follows that:

$$log_{10}(N_0) - 3 = log_{10}(N_0) - log_{10}(1+e^{\frac{log(C_3) - \kappa}{\sigma^2}})$$ Subtracting $log_{10}(N_0)$ and multiplying by -1 gives:

$$log_{10}(1+e^{\frac{log(C_3) - \kappa}{\sigma^2}}) = 3$$ $$1+e^{\frac{log(C_3) - \kappa}{\sigma^2}} = 10^3$$ $$e^{\frac{log(C_3) - \kappa}{\sigma^2}} = 10^3-1$$ $$\frac{log(C_3) - \kappa}{\sigma^2} = log(10^3-1)$$ $$log(C_3) - \kappa = log(10^3-1)\sigma^2$$ $$ \kappa = log(C_3) - log(10^3-1)\sigma^2$$ And substituting this back into the original form of the model gives:

$$log_{10}(N_t) = log_{10}(N_0) - log_{10}(1+e^{\frac{log(x) - (log(C_3) - log(10^3-1)\sigma^2)}{\sigma^2}})$$
